Competing Bandits:\\ The Perils of Exploration under Competition \thanks{This is an extended abstract of \citet{CompetingBandits-merged}, which in turn subsumes conference papers \citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19}.}

### 2020-09-03 journal paper/content/sec-intro.tex
Added some coloring.

The main high-level question is: 
%that we focus on in this paper is:
%\begin{align}\label{eq:main-Q}
\textbf{How does competition incentivize the adoption of better exploration algorithms?}
%\end{align}
This translates into a number of more concrete questions. While it is commonly assumed that better technology always helps, is this so under competition? Does increased competition lead to higher consumer welfare? How significant are ``data feedback loops" --- when more data leads to more users, which leads to even more data, etc. --- and how they relate to the anti-trust considerations? 
%Finding formalizations that admit meaningful answers is a major part of the overall challenge.
We offer a mix of theoretical results and numerical simulations, in which we study complex interactions between platforms' learning dynamics and users' self-interested behavior.

@@

%We investigate these questions with
\xhdr{Our model: competition game.} We consider a stylized duopoly model in which two firms (\emph{principals}) simultaneously engage in exploration and compete for users (\emph{agents}). Principals compete on quality rather than on prices. Agents arrive sequentially, and choose among the two principals according to some decision rule. The chosen principal chooses an action (\eg a list of web search results to show to the user), the user experiences this action, and reports a reward. Each principal only observes its own users. Principals commit to their strategies simultaneously and in advance, with objective to maximize the market share. We investigate several model variants within this framework, where we vary agents' decision rule and/or give a first-mover advantage to one of the principals. In all variants, agents have little or no information about other agents' choices and rewards.

@@

Each principal faces a basic and well-studied version of the multi-armed bandit problem, where each reward is drawn independently from a fixed, action-specific distribution. Each principal's strategy is a bandit algorithm; as such, it adjusts itself to the observed rewards, but not to the competition. This modeling choice reflects the reality of industrial applications, which build on a huge body of knowledge in machine learning. Responding to the competition is another layer of complexity which has not been previously studied in this context, let alone made practical. We draw on the machine learning literature to compare bandit algorithms to one another in the absence of competition, so that we can talk about \emph{better} bandit algorithms in a principled way.%
\footnote{Comparisons between bandit algorithms are somewhat subtle, as some algorithms may be better for some problem instances and/or time intervals, and worse for some others. Generally,  ``better" algorithms are better in the long run, but could be worse initially. More on this in Preliminaries.} One baseline is algorithms that do not explicitly explore (\emph{greedy algorithms}); in the absence of competition, they are known to be terrible for a wide variety of problem instances.

>>>>

The main high-level question is: 
%that we focus on in this paper is:
%\begin{align}\label{eq:main-Q}
\textbf{How does competition incentivize the adoption of better exploration algorithms?}
%\end{align}
This translates into a number of more concrete questions. While it is commonly assumed that better technology always helps, is this so under competition? Does increased competition lead to higher consumer welfare? How significant are ``data feedback loops" --- when more data leads to more users, which leads to even more data, etc. --- and how they relate to the anti-trust considerations? 
%Finding formalizations that admit meaningful answers is a major part of the overall challenge.
\asedit{We offer a mix of theoretical results and numerical simulations, in which we study complex interactions between platforms' learning dynamics and users' self-interested behavior.}

@@

%We investigate these questions with
\asedit{
\xhdr{Our model: competition game.} We consider a stylized duopoly model in which two firms (\emph{principals}) simultaneously engage in exploration and compete for users (\emph{agents}). Principals compete on quality rather than on prices. Agents arrive sequentially, and choose among the two principals according to some decision rule. The chosen principal chooses an action (\eg a list of web search results to show to the user), the user experiences this action, and reports a reward. Each principal only observes its own users. Principals commit to their strategies simultaneously and in advance, with objective to maximize the market share. We investigate several model variants within this framework, where we vary agents' decision rule and/or give a first-mover advantage to one of the principals. In all variants, agents have little or no information about other agents' choices and rewards.

Each principal faces a basic and well-studied version of the multi-armed bandit problem, where each reward is drawn independently from a fixed, action-specific distribution. Each principal's strategy is a bandit algorithm; as such, it adjusts itself to the observed rewards, but not to the competition. This modeling choice reflects the reality of industrial applications, which build on a huge body of knowledge in machine learning. Responding to the competition is another layer of complexity which has not been previously studied in this context, let alone made practical. We draw on the machine learning literature to compare bandit algorithms to one another in the absence of competition, so that we can talk about \emph{better} bandit algorithms in a principled way.%
\footnote{Comparisons between bandit algorithms are somewhat subtle, as some algorithms may be better for some problem instances and/or time intervals, and worse for some others. Generally,  ``better" algorithms are better in the long run, but could be worse initially. More on this in Preliminaries.} One baseline is algorithms that do not explicitly explore (\emph{greedy algorithms}); in the absence of competition, they are known to be terrible for a wide variety of problem instances.}

@@
